---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# ggex

<!-- badges: start -->
<!-- badges: end -->

ggex (think 'FedEx') is for 'ex' as in fast and expressive and 'ex' for extension/extenders.


## `qlayer()`, is `layer()` with plenty of defaults

```{r qlayer}
qlayer <- function (mapping = NULL, data = NULL, geom = ggplot2::GeomPoint, 
                    stat = ggplot2::StatIdentity, 
    position = position_identity(), ..., na.rm = FALSE, show.legend = NA, 
    inherit.aes = TRUE) 
{
    ggplot2::layer(data = data, mapping = mapping, geom = geom, 
        stat = stat, position = position, show.legend = show.legend, 
        inherit.aes = inherit.aes, params = rlang::list2(na.rm = na.rm, 
            ...))
}
```

```{r}
library(ggplot2)
ggplot(data = cars) + 
  aes(speed, dist) + 
  qlayer()

last_plot() + 
  qlayer(geom = GeomTile)

last_plot() + 
  qlayer(position = "jitter", color = "red")

last_plot() + 
  ggchalkboard::theme_chalkboard()

```

## `combine_aes()` to make small adjustments to exising `default_aes` settings

```{r}
combine_aes <- utils::modifyList
```


```{r}
GeomPoint$default_aes

combine_aes(GeomPoint$default_aes, aes(color = "blue"))


GeomBluePoint <- ggproto("GeomBluePoint", 
                         GeomPoint, 
                         default_aes = 
                           combine_aes(GeomPoint$default_aes, 
                                       aes(color = "blue")))

ggplot(cars) + 
  aes(speed, dist) + 
  stat_identity(geom = GeomBluePoint)
```



### `combine_aes()` examples containing recommended usage for new geom theming... use `from_theme()`, `col_mix()`, `paper`, `ink`, `accent` etc.

#### GeomBubble, GeomBubbleColorMixFilled, GeomPointToAccentColor

```{r}
GeomPoint$default_aes

GeomBubble <- ggproto("GeomBubble", 
                      GeomPoint, 
                      default_aes = combine_aes(GeomPoint$default_aes, 
                                                aes(shape = 21,
                                                    size = from_theme(pointsize*4))))

GeomBubble$default_aes

ggplot(cars) + 
  aes(speed, dist) + 
  qlayer(geom = GeomBubble)

aes_new_bubble_colmix <- aes(shape = 21,
                             size = from_theme(pointsize*4),
                             fill = from_theme(ggplot2:::col_mix(ink, paper, 0.35)))

GeomBubbleColorMixFilled <- ggproto("GeomBubbleColorMixFilled", 
                              GeomPoint, 
                              default_aes = combine_aes(GeomPoint$default_aes,
                                                        aes_new_bubble_colmix))

ggplot(cars) + 
  aes(speed, dist) + 
  qlayer(geom = GeomBubbleColorMixFilled)

GeomPointToAccentColor <- ggproto("GeomPointToAccentColor",  
                                GeomPoint,
                                default_aes = combine_aes(GeomPoint$default_aes,
                                                          aes(color = from_theme(accent))))

ggplot(cars) + 
  aes(speed, dist) + 
  qlayer(geom = GeomPointToAccentColor, size = 6, alpha = .5) + 
  geom_smooth()

last_plot() + 
  ggchalkboard::theme_chalkboard()

new_accent <- ggplot2:::col_mix("purple", "lightyellow", .7)
new_accent

last_plot() + 
  ggchalkboard::theme_chalkboard(accent = new_accent)
```


#### GeomPolygonWithOutline, GeomPolygonOutline

```{r}
GeomPolygonWithOutline <- ggproto("GeomPolygonWithOutline", 
                                  GeomPolygon,
                                  default_aes = combine_aes(GeomPolygon$default_aes,
                                                            aes(color = from_theme(ink),
                                                                fill = from_theme(ggplot2:::col_mix(ink, paper, 0.85)))))
  
GeomPolygonOutline <- ggproto("GeomPolygonOutline",  # New Geom
                              GeomPolygon, # inherit
                              default_aes = combine_aes(GeomPolygon$default_aes,
                                                        aes(color = from_theme(ink),
                                                            fill = NA)))

ggplot(cars) + 
  aes(speed, dist) + 
  qlayer(geom = GeomPolygonWithOutline)

last_plot() + 
  ggchalkboard::theme_chalkboard()

ggplot(cars) + 
  aes(speed, dist) + 
  qlayer(geom = GeomPolygonOutline)

last_plot() + 
  ggchalkboard::theme_chalkboard()
```


#### GeomLabelClearBackground, GeomLabelWithPanelBackground, GeomLabelWithPanelBackground

```{r}
GeomLabelClearBackground <- ggproto("GeomLabelClearBackground",
                                    GeomLabel,
                                    default_aes = 
                                      combine_aes(GeomLabel$default_aes, 
                                                  aes(fill = NA, linewidth = 0)))

ggplot(cars) + 
  aes(speed, dist, label = dist) +
  geom_point() +
  qlayer(geom = GeomLabelClearBackground, 
         vjust = "outward",
         hjust = "outward")

last_plot() + 
  ggchalkboard::theme_chalkboard()



# I'm pretty sure this isn't right, and maybe 'right' is not possible, but I like the idea of it.
GeomLabelWithPanelBackground <- ggproto("GeomLabelWithPanelBackground",
                                    GeomLabel,
                                    default_aes = 
                                      combine_aes(GeomLabel$default_aes, 
                                                  aes(fill =
                                                        from_theme(ggplot2:::col_mix(ink, paper, .92)),
                                                      linewidth = 0)))
                                    
ggplot(cars) + 
  aes(speed, dist, label = dist) +
  geom_point() +
  qlayer(geom = GeomLabelWithPanelBackground, 
         vjust = "outward",
         hjust = "outward"
         )


# also not 100% sure what's going on here - why does label outline not disappear?
GeomLabelRepelWithPanelBackground <- ggproto("GeomLabelRepelWithPanelBackground",
                                    ggrepel::GeomLabelRepel,
                                    default_aes = 
                                      combine_aes(ggrepel::GeomLabelRepel$default_aes, 
                                                  aes(fill =
                                                        from_theme(ggplot2:::col_mix(ink, paper, .92)),
                                                      label.size = 0,
                                                      linewidth = 0)))


ggplot(cars) + 
  aes(speed, dist, label = dist) +
  geom_point() +
  qlayer(geom = GeomLabelRepelWithPanelBackground)

```

## `qproto_update()` lets you use an existing proto object, minimally modified, in a layer(), qlayer(), geom_\*() or stat_*() function.  


```{r}
proto_update <- function (`_class`, `_inherit`, default_aes_update = NULL, ...) {
    if (!is.null(default_aes_update)) {
        
      new_default_aes <- combine_aes(`_inherit`$default_aes, default_aes_update)
      
    } else {
      
      new_default_aes <- `_inherit`$default_aes
      
    }
  
    ggplot2::ggproto(`_class` = `_class`, 
                     `_inherit` = `_inherit`,
                     default_aes = new_default_aes, ...)
}


qproto_update <- function (`_inherit`, default_aes_update = NULL, ...){
    proto_update("protoTemp", `_inherit`, default_aes_update = default_aes_update, 
        ...)
}
```


```{r}
ggplot(cars) + 
  aes(speed, dist) + 
  qlayer(geom = qproto_update(GeomPoint, aes(shape = 21)), # try on geom w modified default aes
         size = 8)
```


## The `qstat()` family lets you try out Stats without naming them.


```{r}
qstat <- function (compute_group = ggplot2::Stat$compute_group, ...) {
    ggplot2::ggproto("StatTemp", Stat, compute_group = compute_group, 
        ...)
}
```


```{r}
compute_group_mean <- function(data, scales){
  
  data |> dplyr::summarise(x = mean(x), y = mean(y))
  
}

ggplot(cars) + 
  aes(speed, dist) + 
  geom_point() +
  qlayer(geom = qproto_update(GeomPoint, 
                              aes(size = from_theme(pointsize * 3))),
         stat = qstat(compute_group_mean))

last_plot() + 
  ggchalkboard::theme_chalkboard()
```


```{r}
qstat_group <- qstat

qstat_layer <- function (compute_layer, ...){
    ggplot2::ggproto("StatTemp", Stat, compute_layer = compute_layer, 
        ...)
}

qstat_panel <- function (compute_panel, ...){
    ggplot2::ggproto("StatTemp", Stat, compute_panel = compute_panel, 
        ...)
}
```




## `vars_pack()` and `vars_unpack()` allow for multi-var aesthetic mapping.

June wrote this...

https://github.com/ggplot2-extenders/ggplot-extension-club/discussions/18#discussioncomment-10219152


> The major challenge here is that you'd want a single column to represent values across all of the predictors, so you'd need to do some kind of compression of values to make them fit inside a single column in the layer data. At least for PCA the predictors are of the same type, so you could make predictors a list-column of length-3 numeric vectors.

> I edited your code a bit, introducing the nesting-unnesting functions vars_pack() and vars_unpack(). Essentially, the predictor columns are nested into a single column and lays dormant until you need it for PCA, at which point it's unnested back into a dataframe. Also note that remove_missing() should happen after the unnesting to spot individual NA values.


```{r}
library(tidyverse)

vars_pack <- function(...) {
  
  varnames <- as.character(ensyms(...))
  vars <- list(...)
  listvec <- asplit(do.call(cbind, vars), 1)
  structure(listvec, varnames = varnames)

}


palmerpenguins::penguins |> 
  mutate(outcome = species, 
         predictors = vars_pack(bill_length_mm, bill_depth_mm, flipper_length_mm)) |> 
  select(outcome, predictors) ->
data_packed

head(data_packed)

vars_unpack <- function(x) {
  pca_vars <- x
  df <- do.call(rbind, pca_vars)
  colnames(df) <- attr(pca_vars, "varnames")
  as.data.frame(df)
  
}


data_packed %>%
    mutate(vars_unpack(predictors)) %>% 
  select(-predictors) ->
data_uppacked

head(data_uppacked)

```


```{r}

compute_lm_multi <- function(data, scales){
  
  data %>% 
    remove_missing() %>% 
    select(x,y, predictors) %>% 
    mutate(vars_unpack(predictors)) %>% 
    select(-predictors) ->
  data
  
  lm <- lm(data = data, formula = y ~ .)
  
  data$yend = data$y
  data$y = lm$fitted.values
  data$xend = data$x

  data$residuals <- lm$residuals
  
  data
}

variables <- vars_pack

palmerpenguins::penguins %>% 
  remove_missing() %>% 
  mutate(x = bill_length_mm, y = bill_depth_mm, predictors = variables(sex, species)) %>% 
  select(x, y, predictors) %>% 
  compute_lm_multi()
 
palmerpenguins::penguins |>
  remove_missing() |>
  ggplot() + 
  aes(y = flipper_length_mm, x = bill_length_mm, predictors = variables(species)) + 
  geom_point() + 
  qlayer(stat = qstat_panel(compute_lm_multi), 
         geom = qproto_update(GeomPoint, 
                              aes(color = from_theme(accent)))) + 
  qlayer(stat = qstat_panel(compute_lm_multi), 
         geom = qproto_update(GeomSegment,
                              aes(color = from_theme(accent))),
         linetype = "dotted")

last_plot() + 
  geom_smooth(color = "red", method = lm) + 
  aes(group = species)

last_plot() + 
  ggchalkboard::theme_chalkboard()
```

---

# heat map experiment with https://github.com/teunbrand/ggplot_tricks

```{r}

# seems like the greys might be switched out with ink and paper? Though maybe not...
# from https://github.com/teunbrand/ggplot_tricks
contrast <- function(colour) {
  out   <- rep("grey20", length(colour))
  light <- farver::get_channel(colour, "l", space = "hcl")
  out[light < 50] <- "grey80"
  out
}

# from https://github.com/teunbrand/ggplot_tricks
aes_autocontrast <- aes(colour = after_scale(contrast(fill)))

tidy_titanic <- Titanic |> data.frame() |> tidyr::uncount(Freq)
tidy_titanic |> head()

ggplot(tidy_titanic) + 
  aes(Sex, Survived) + 
  # sketch of geom_heat
  qlayer(geom = GeomTile,
         stat = qproto_update(StatSum, 
                              aes(fill = after_stat(n),
                                  size = NULL))) + 
  # sketch of geom_heat_text
  qlayer(geom = qproto_update(GeomText, aes_autocontrast),
         stat = qproto_update(StatSum,
                              aes(label = after_stat(n),
                                  fill = after_stat(n),
                                  size = NULL,
                                  )))

last_plot() + 
  ggchalkboard::theme_chalkboard()


```


